---
title: "TCRBroom: an example workflow"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false

vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# 1. Introduction
The `TCRBroom` package is designed to check pairwise overlap and cross-cohort overlap
for TCR samples. This vignette shows a simple workflow illustrating how TCRBroom package works. 


## Installation
Before the installation, please make sure you have installed JAGS-4.x.y.exe from http://www.sourceforge.net/projects/mcmc-jags/files. 

The package can be installed using `devtools` by the following commands
```{r, results=FALSE, warning=FALSE, message=FALSE}
library(devtools)
install_github("rx-li/TCRBroom")
```

After the installation, the package can be loaded with
```{r, results=FALSE, warning=FALSE, message=FALSE}
library(TCRBroom)
```

# 2. Example workflow

## Plot pairwise and crosscohort overlaps 
The TCR-seq data and meta table that we used in this example workflow can be found under the folder `data/`. 
18 samples from 10 patients are simulated based on the real contaminated data. 

First, we need to read the meta table, which documents sample names and 
their corresponding subject IDs. 
```{r}
meta <- read.csv("../data/meta.csv")
# Please make sure that the column name are exactly "sample_name" and "subject_id".
colnames(meta) <- c("sample_name", "subject_id")
head(meta)
dim(meta)
```

### Pairwise overlap
We can check the pairwise overlap between each two subjects using the following command. 
```{r}
pairwise_table <- checkPairwise(meta, "../data/")
```

The heatmap shows the magnitude of the pairwise overlaps. 
A matrix is also created for users to check the values of the overlaps:
```{r}
pairwise_table[1:3, 1:3]
```

### Cross-cohort overlap
We can call the function `checkCrosscohort` to check the cross-cohort overlap (sequence prevalence). 
```{r, warning=FALSE, message=FALSE}
crosscohort_table <- checkCrosscohort(meta, "../data/", seq.n=30)
```

By default, the function will generate an heatmap showing prevalence for the 30 most prevalent sequences.

### Samples with few seqeunces identification
If users would like to identify the subjects with relatively fewer sequences (such as subjects with less than 3000 sequences), `ssampleIndicator` can generate a table to note those subjects. 
```{r}
seq_summary <- ssampleIndicator(meta, "../data/", low.seq=3000)
head(seq_summary)
```
Subjects with sequences fewer than 3,000 will be flagged as "Alert" in the "low_seq" column.

## Identify outliers

### Pairwise overlap
In order to identify pairwise overlap outliers, user need to provide a hierarchical `jags` model 
and the initial values of the parameters for the model. 
```{r}
model <- function(){
  for(k in 1:k){
    y[k]~dbeta(alpha, beta) 
  }
  # Specify the prior based on the clean data
  alpha~dlnorm(1.14, 0.51)
  beta~dlnorm(9.60, 1.09)
}
model_params <- c("alpha", "beta")
model_inits <- function(){
  list("alpha" = 1, "beta" = 1000) # Initial value of the parameteres
}
```

Another step is to format the pairwise overlap table as a named vector.
```{r}
# Extract the subject's names
names_vector <- colnames(pairwise_table)
names_paired <- unlist(
  lapply(c(2:ncol(pairwise_table)), function(x) paste0(names_vector[x], "+", names_vector[1:x-1]))
  )
pairwise_v <- as.vector(pairwise_table)
pairwise_v <- na.omit(pairwise_v)
names(pairwise_v) <- names_paired
# We used a beta distribution in the example, so we will filter out 0s in the vector. 
pairwise_v <- pairwise_v[pairwise_v > 0]
```

Then, function `detectOutliers` can be called to identify the outliers given 
a cutoff value.  
```{r}
results <- detectOutliers(pairwise_v, iters=10, c=0.95, model_inits, model_params, model)
```


After 3 rounds, all outliers were detected. We can take a look at the output.
```{r}
str(results)
```
The output includes the corresponding critical value of the cutoff, outliers, and
the parameters of the posterior distribution for each round. 

We can check the outliers by 
```{r}
# Outliers are detected after 3 rounds
results$outliers[[3]]
```
The first outlier is the overlap between subjects "P04" and "P02". 
We can check the overlap further using function `check2Samples`. 
```{r}
tables_p04_p02 <- check2Samples(meta, "P02", "P06", data_dir="../data/")
```

"tables_p04_p02" will include two matrix: One show the number of overlap between 
any two samples of subjects "P04" and "P02". The other show the percentage of overlap 
between any two samples of subjects "P04" and "P02".
```{r}
tables_p04_p02$overlap_n
tables_p04_p02$overlap_pct
```
Moreover, we can visualize the percentage of the overlap to get a better sense of which sample 
can be thought as a source of contamination. 
```{r}
library(RColorBrewer)
library(gplots)
Colors=rev(brewer.pal(11,"Spectral"))
Colors=colorRampPalette(Colors)(100)

table_pct <- tables_p04_p02$overlap_pct
# rename the table for visualization
colnames(table_pct) <- gsub(".tsv", "", colnames(table_pct))
rownames(table_pct) <- gsub(".tsv", "", rownames(table_pct))
# create the heatmap
heatmap.2(as.matrix(table_pct[1:nrow(table_pct)-1, 1:ncol(table_pct)-1]), 
          col=Colors, Rowv=FALSE, Colv=FALSE,
          dendrogram='none', trace="none", cexRow=1, cexCol=1, density.info="none")
```

We can see that sample "s2_2" have relatively heavy overlap with the samples of subject "P06".
We can consider to remove sample "s2_2". 

### Cross-cohort overlap
Empirical P values can be calculated for the cross-cohort overlap values using a reference vector. 
For this example, we will randomly generate a reference vector. 
```{r}
# Randomly generate the reference vector 
ref_v <- runif(50000, 0, 0.3)
# extract the prevalence proportion
crosscohort_v <- crosscohort_table[, dim(crosscohort_table)[2]]
# calculate the empirical P values
p_table <- Pcalculator(ref_v, crosscohort_v)
# Here we eliminate row names for presenting since the sequences' names are extremely long.
rownames(p_table) <- NULL
head(p_table)
```
If we set the cutoff value for adjusted P value as 0.05, then we can extract the outliers by
```{r}
outliers <- p_table[p_table$p_adjust <= 0.05, ]
```



```{r}
sessionInfo()
```





