---
title: "TCRBroom: an example workflow"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false

vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 1. Introduction
The `TCRBroom` package is designed to check pairwise overlap and cross-cohort overlap
for TCR samples. This vignette shows a simple workflow illustrating how TCRBroom package works. 


### Installation
Before the installation, please make sure you have installed JAGS-4.x.y.exe from http://www.sourceforge.net/projects/mcmc-jags/files. 

The package can be installed using `devtools` by the following commands
```{r, results=FALSE, warning=FALSE, message=FALSE}
library(devtools)
install_github("rx-li/TCRBroom")
```

After the installation, the package can be loaded with
```{r, results=FALSE, warning=FALSE, message=FALSE}
library(TCRBroom)
```

## 2. Example workflow

### Plot pairwise and crosscohort overlaps 
The data we used in this example workflow can be found in the folder `data/`. 18 samples from 10 patients are simulated based on the real contaminated data. 

First, we need to read the meta files. The meta files includes sample names and 
their corresponding subject ID. 

```{r}
meta <- read.csv("../data/meta.csv")
colnames(meta) <- c("sample_name", "subject_id")
head(meta)
dim(meta)
```

Then, we can check the pairwise overlap using the following command. 
```{r}
pairwise_table <- checkPairwise(meta, "../data/")
```
The heatmap shows the magnitude of the pairwise overlaps. 
A matrix is also created for users to check the values of the overlaps:
```{r}
pairwise_table[1:3, 1:3]
```

We can also call the function to check the cross-cohort overlap (sequence prevalence). 
```{r, warning=FALSE, message=FALSE}
crosscohort_table <- checkCrosscohort(meta, "../data/", seq.n=30)
```
By default, the function will generate an OncoPrint showing prevalence for the 30 most prevalent sequences.

If users would like to identify the subjects with relatively fewer sequences (such as subjects with less than 3000 sequences), `ssampleIndicator` can generate a table to note those subjects. 

```{r}
seq_summary <- ssampleIndicator(meta, "../data/", low.seq=3000)
head(seq_summary)
```
Subjects with sequences fewer than 3,000 will be noted as "Alert" in the "low_seq" column.

### Identify outliers
In order to identify pairwise overlap outliers, user need to provide a hierarchical model and initials of the parameters for the model.
```{r}
model <- function(){
  for(k in 1:k){
    y[k]~dbeta(alpha, beta)
  }
  alpha~dlnorm(1.14, 0.51)
  beta~dlnorm(9.60, 1.09)
  # specify the prior based on the clean data
  # get the alpha and beta
  # find the outlier iterately 
}
model_params <- c("alpha", "beta")
model_inits <- function(){
  list("alpha" = 1, "beta" = 1)
}
```

Then the pairwise overlap values need to be formatted as a named vector.
```{r}
names_vector <- colnames(pairwise_table)
names_paired <- unlist(
  lapply(c(2:ncol(pairwise_table)), function(x) paste0(names_vector[x], "+", names_vector[1:x-1]))
  )
pairwise_v <- as.vector(pairwise_table)
pairwise_v <- na.omit(pairwise_v)
names(pairwise_v) <- names_paired
# We used a Gamma model in the example, so we will filter out 0s in the vector. 
pairwise_v <- pairwise_v[pairwise_v > 0]
```

After that, a function can be called to identify the outliers given a quantile cutoff. 
```{r}
results <- detectOutliers(pairwise_v, iters=10, c=0.95, model_inits, model_params, model)
```
At the 3rd iteration, all outliers were detected. We can check the outliers by 
```{r}
results$outliers[[3]]
```
Empirical P values can be calculated for the cross-cohort overlap values using a reference vector. 
For this example, we randomly generate a reference vector. 
```{r}
ref_v <- runif(50000, 0, 0.3)
# extract the prevalence proportion
crosscohort_v <- crosscohort_table[, dim(crosscohort_table)[2]]
# calculate the empirical P values
p_table <- Pcalculator(ref_v, crosscohort_v)
# Here we eliminate row names for presenting since the sequences' names are extremely long.
rownames(p_table) <- NULL
head(p_table)
```




```{r}
sessionInfo()
```





